---
title: "LA Demographics"
author: "Julia Ma"
output: html_document
---

Mapping LA Project is useful if you'd like to view demographics given a region/neighborhood. However, what if you'd like to view neighborhoods given demographics? The drop-down menu from http://maps.latimes.com/neighborhoods/ works, however is difficult to navigate. Goal: implement dynamic visualization using RShiny to filter by demographic sections then yield regions/neighborhoods.

# Setting up baseline - Region and Neighborhoods

```{r}
library(rvest)
library(dplyr)

# import LA neighborhoods and regions
mainn <- read_html("http://maps.latimes.com/neighborhoods/neighborhood/list/")

# grab links of neighborhoods and regions, format df
links <- mainn %>% 
  html_nodes(".datagrid a") %>% 
  html_attr("href")

links <- paste0("http://maps.latimes.com", links)
links <- as.data.frame(matrix(links, ncol = 2, byrow = T), 
                       stringsAsFactors = F)
names(links) <- c("Neighborhood", "Region")

# str(links)
n_neighborhood <- length(unique(links[,1])) #272 unique neighborhoods

```


# Demographics

## Population - inefficient?

```{r}
# #grab population statistics
# 
# neighborhoods <- data.frame(NULL)
# 
# fcn_popu <- function(website){
#   website <- read_html(website)
#   
#   name <- website %>% 
#     html_nodes("h1") %>% 
#     html_text("href")
# 
#   population <- website %>%
#     html_nodes(":nth-child(4) li:nth-child(1) strong") %>%
#     html_text("href") %>%
#     gsub(",", "", .) %>%
#     as.numeric(.)
# 
#   sqmi <- website %>%
#     html_nodes(":nth-child(4) li:nth-child(2) strong") %>%
#     html_text("href") %>%
#     as.numeric(.)
# 
#   density <- round(population/sqmi, 2)
# 
#   return(c(name, population, sqmi, density))
# }
# 
# 
# demo_popu_raw <- as.data.frame(do.call(rbind, lapply(links[,1], fcn_popu)), ncol = 4)
# 
# # format df
# cols_need_num <- 2:4
# demo_popu <- demo_popu_raw[,1:4]
# # apply(levels(demo_popu_raw[,2:4])[demo_popu_raw[,2:4]], as.numeric) -- couldn't use apply over mutiple cols??
# 
# demo_popu[,1] <- as.character(demo_popu_raw[,1])
# demo_popu[,2] <- levels(demo_popu_raw[,2])[demo_popu_raw[,2]]
# demo_popu[,3] <- levels(demo_popu_raw[,3])[demo_popu_raw[,3]]
# demo_popu[,4] <- levels(demo_popu_raw[,4])[demo_popu_raw[,4]]
# 
# names(demo_popu) <- c("Name", "Population (2000)", "Sq Mi", "Density")
# 
# #some rows have duplicated names across all columns, remove those duplicates
# duplicates <- which(demo_popu[,1] == demo_popu[,2])
# demo_popu[duplicates, 2:4] <- NA
# 
# #View webscraped data
# 
# # summary(demo_popu) many NA's
# # View(demo_popu[demo_popu[,1] %in% demo_popu[,2],]) have NA data
```


# Region and Demographics - Efficient?? all demo at once

###  Initiate df with column of names

```{r}
df <- read_html("http://maps.latimes.com/neighborhoods/neighborhood/list/") %>% 
  html_nodes("td:nth-child(1) a") %>% 
  html_text("href")
df <- df[-273] # drop extra last row
df <- as.data.frame(df, ncol = 1)
names(df) <- "Neighborhood"
```

### obtain demographics options from external file
```{r}
all <- as.character(unlist(read.csv("urllist.txt", header = F)))
demographics_keyword <- c("/list/")
demographics_links <- all[grepl(paste(demographics_keyword, collapse = "|"), all)]
#remove first extra element
demographics_links <- demographics_links[-1] 

#fill in missing links unavailable from the sitemap
demographics_links <- c(demographics_links,                        "http://maps.latimes.com/neighborhoods/owners/neighborhood/list/", "http://maps.latimes.com/neighborhoods/renters/neighborhood/list/",
"http://maps.latimes.com/neighborhoods/area/square-miles/neighborhood/list/", 
"http://maps.latimes.com/neighborhoods/ethnicity/non-white/neighborhood/list/", 
"http://maps.latimes.com/neighborhoods/population/total/neighborhood/list/")

demographics_links <- demographics_links[order(demographics_links)]

# number of demographic links is 48, which matches the drop-down-bar
```

### import links' data and clean
```{r}
clean_demo <- function(demo){
  cleaned <- demo[,c(2:3)]
  
  #convert % to numeric type
  if(grepl("%", cleaned[1,2])){
    names(cleaned)[2] <- paste(names(cleaned)[2], "(%)")  
    cleaned[,2] <- as.numeric(gsub("%", "", cleaned[,2])) / 100
  }
    
  #convert $ to numeric type
  if(grepl("\\$", cleaned[1,2])){
    names(cleaned)[2] <- paste(names(cleaned)[2], "($)")  
    cleaned[,2] <- as.numeric(gsub("[[:punct:]]", "", cleaned[,2]))
  }
  
  return(cleaned)
}
```  

### merge all the demographics info together
```{r}
library(dplyr)

for(i in seq_along(demographics_links)){
  # construct dataframe for each demographic option
  one_demo <- (read_html(demographics_links[i]) %>% 
    html_nodes("table") %>% 
    html_table())[[2]]
  
  cleaned_demo <- clean_demo(one_demo)
  
  df <- left_join(df, cleaned_demo, by = "Neighborhood")
}

#rename cols 24:27
names(df)[24:27] <- c("20k income or less percent of households", 
                      "20k to 40k income percent of households", 
                      "40k to 60k income percent of households", 
                      "60k to 125k income percent of households")
  
```

# data exploration
```{r}
# summary(df)
# lapply(df, function(x) which(is.na(x)))
# therefore remove corresponding rows

# dim(df)
# dim(df[-unlist((lapply(df, function(x) which(is.na(x)))[2])),])

df_clean <- df[-unlist((lapply(df, function(x) which(is.na(x)))[2])),]

hist(df$`Age 10 or less (%)`)
```

#shiny? to be continued. . .. 
```{r}

```
